---
title: "Mini Project 3: Trail Tinkering"
author: "Lucy Hall and Lauren Low"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Problem Statement
MacLeish Field Station, in Whately, Massachusetts, is a Smith owned property a little under 30 minutes north east of Smith College.  MacLeish is home to four miles of hiking trails, available to all visitors and guests.  As of now, the trails do not have a difficulty classification.  The task at hand is to create a classification of MacLeish trails (easy, moderate, or difficult) based on length, change in elevation, and terrain.[^1]

[^1]:Github repository: https://github.com/spf500/macleish

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(macleish)
library(leaflet)
library(sf)
library(knitr)
library(RColorBrewer)
```

We created a table called Macleish_trails that displays the trails within macleish_layers.
```{r, message = FALSE, warning = FALSE}
Macleish_trails <- macleish_layers %>%
  pluck("trails")

Macleish_trails
```

We added in contours by plucking the countours_3m data from the orginal dataset
```{r, message = FALSE, warning = FALSE}
Contours <- macleish_layers %>%
  pluck("contours_3m")

Contours
```

We modified Macleish_trails (and called the modification length_trails) so that the table would display the length of each trail.
```{r, message = FALSE, warning = FALSE}
length_trails <-Macleish_trails %>%
  mutate(length = st_length(geometry)) %>%
  group_by(name) %>%
  summarise(total_length = sum(length))%>%
  arrange(desc(total_length))

length_trails
```

We made the dataframe look a little nicer.
```{r, message = FALSE, warning = FALSE}
  length_trails_df <- length_trails %>%
  st_set_geometry(NULL)
knitr::kable(length_trails_df)
```

We created a map that categorized the trails by length, according to color.
```{r, message= FALSE, warning = FALSE}
length_colors <- colorQuantile("YlOrRd", length_trails %>%
                              pull(total_length))

length_colors
```

```{r, message = FALSE, warning = FALSE}
rankings_map <- leaflet() %>%
  addTiles %>%
  addPolylines(data = length_trails, color = ~length_colors(total_length), weight = 2, opacity = 1) %>%
  addPolylines(data = Contours, weight = 1, opacity = 0.2, options = pathOptions(interactive = FALSE)) %>%
  addLegend(pal = length_colors, values = length_trails$total_length) 

rankings_map
```

We found the points where the contrours and trails intersect
```{r, message = FALSE, warning = FALSE}
intersections <- Macleish_trails %>%
  st_intersection(Contours)

intersections
```

We matched trails to colors, by trail names.
```{r, message= FALSE, warning = FALSE}
trail_colors <- colorFactor("Set1", Macleish_trails %>%
                              pull(name))
trail_colors
```

We created a map that plotted the trails at MacLeish, each trail being a different color.
```{r, message = FALSE, warning = FALSE}
trail_map <- leaflet() %>%
  addTiles %>%
  addPolylines(data = Contours, weight = 1, opacity = .3) %>%
  addPolylines(data = Macleish_trails, color = ~trail_colors(name), weight = 2, opacity = 1) %>%
  addLegend(pal = trail_colors, values = Macleish_trails$name) 

trail_map
```

We modified Macleish_trails to show the difficulty of each trail.
```{r, message = FALSE, warning = FALSE}
Macleish_trails["difficulty"] <- c("moderate","difficult","moderate","easy","moderate","difficult","difficult","easy","moderate","easy","moderate","easy","difficult","difficult","easy")

Macleish_trails
```

We matched trails to colors, by trail difficulty.
```{r, message= FALSE, warning = FALSE}
trail_colors2 <- colorFactor("Set1", Macleish_trails %>%
                              pull(difficulty))
```

We created a map that shows tre trails by difficulty
```{r, message = FALSE, warning = FALSE}
initial_map <- leaflet() %>%
  addTiles %>%
  addPolylines(data = Contours, weight = 1, opacity = .3) %>%
  addPolylines(data = Macleish_trails, color = ~trail_colors2(difficulty), weight = 2, opacity = 1) %>%
  addLegend(pal = trail_colors2, values = Macleish_trails$difficulty) 
initial_map
```

We converted MULTIPOINTs to POINTs.
```{r, message = FALSE, warning = FALSE}
cast_intersections <- intersections %>%
  st_cast("MULTIPOINT")%>%
  st_cast("POINT")

cast_intersections
```

We took a look at the elevation for each trail.
```{r, message = FALSE, warning = FALSE}
all_loops <- cast_intersections %>%
  select("name", "ELEV_M") %>%
  st_set_geometry(NULL)
all_loops
```

We made a function called elevation_profile and applied it to the different trails to look at the elevation along the trail.
```{r, message = FALSE, warning = FALSE}
elevation_profile <- function(trail_name) {
  all_loops %>%
    filter(name == trail_name) %>%
    mutate(id = row_number()) %>%
    ggplot(aes(x = id, y = ELEV_M)) + 
    geom_point() + 
    geom_line() + 
    expand_limits(x = c(0,50), y = c(210, 290)) +
    labs(title = trail_name)
}

knitr::opts_chunk$set(echo = TRUE, fig.show = 'hold', out.width = '33%')

elevation_profile("Porcupine Trail")
elevation_profile("Western Loop")
elevation_profile("Easy Out")
elevation_profile("Eastern Loop")
elevation_profile("Snowmobile Trail")
elevation_profile("Poplar Hill Road")
elevation_profile("entry trail")
elevation_profile("Driveway")
elevation_profile("Vernal Pool Loop")
```

##A Recommendation
We propose that Smith College and MacLeish Field Station design a new map for MacLeish.  This map will be a color-coded map, which will indicate the difficulty of each trail while showing contours for extra indication of difficulty.  We also recommend that MacLeish post signs all throughout each trail (i.e. an easy trail will have a green triangle mark, a moderate trail will have a blue square mark, and a difficult trail will have a black diamond mark).  Each sign will correspond to the colors/symbols on the MacLeish map.

Based on a rough cost analysis, implementing signs and new maps would cost about $300.  There are nine trails, and we would like to put a sign at the beginning and end of each trail, which results in the creation of 18 signs.  Each sign will cost about $10 to create, thus creating trail signs for all of MacLeish will cost $180.  Further, we hope to purchase 4 gallons of paint (green, blue, black, and brown) to paint signs as well as paint blazes on trees.  Each gallon will cost about $30, resulting in a total cost of $120 for paint

Developing these signs as blazes should take no more than one week, especially if the weather is cooperative.  Signs can be developed and created anywhere with access to power tools, including  hand drills and saws.  On a sunny day, blazes should be painted on to rocks and trees along each trail.  It is also important that the following day be sunny, as the rain has the potential to remove undried paint.  Many hands should be on deck for this project.  For optimal results, it would be beneficial to have a team of 10-20 people, one to two people per trail working on painting blazes and five plus people working on creating/distributing signs.

There should be few environmental impacts of this project.  However, it is important that we make sure to paint blazes around animal habitats, rather than on them (i.e. paint a blaze on a tree trunk rather than near a hole/nest to avoid habitat disturbance.  Further, there are few road blocks other than determining where to place blazes and signs along the trail.